\documentclass{openreader}
\title{Ἰωάννου γʹ}
\date{}
\setmainlanguage{english}
\setmainfont{Charis SIL}
\setotherlanguage{greek}
\newfontfamily\greekfont[Script=Greek]{SBL BibLit}
\begin{document}
\selectlanguage{greek}
\maketitle
\raggedbottom 
\fontsize{16pt}{24pt}\selectfont
\section*{1}
\FootnoteHeader{Chapter 1}
\VerseFootnote{1}{\FootnoteSubHeader{a}\FnParseFormGloss{dative masculine singular}{Γάϊος}{Gaius}\FootnoteSubFooter\FootnoteSubHeader{b}\FnParse{present active indicative 1st singular}\FootnoteSubFooter}\MainTextVerseMark{1}{1}Ὁ πρεσβύτερος Γαΐῳ\MainTextFootnoteMark{a} τῷ ἀγαπητῷ, ὃν ἐγὼ ἀγαπῶ\MainTextFootnoteMark{b} ἐν ἀληθείᾳ. \VerseFootnote{2}{\FootnoteSubHeader{a}\FnParseFormGloss{present middle indicative 1st singular}{εὔχομαι}{to pray for}\FootnoteSubFooter\FootnoteSubHeader{b,d}\FnParseFormGloss{present passive infinitive}{εὐοδόομαι}{to get along with}\FootnoteSubFooter\FootnoteSubHeader{c}\FnParseFormGloss{present active infinitive}{ὑγιαίνω}{to be healthy}\FootnoteSubFooter}\MainTextVerseMark{1}{2}Ἀγαπητέ, περὶ πάντων εὔχομαί\MainTextFootnoteMark{a} σε εὐοδοῦσθαι\MainTextFootnoteMark{b} καὶ ὑγιαίνειν,\MainTextFootnoteMark{c} καθὼς εὐοδοῦταί\MainTextFootnoteMark{d} σου ἡ ψυχή. \VerseFootnote{3}{\FootnoteSubHeader{a}\FnParse{aorist passive indicative 1st singular}\FootnoteSubFooter\FootnoteSubHeader{b}\FnFormGloss{λίαν}{very much}\FootnoteSubFooter\FootnoteSubHeader{c}\FnParse{present middle participle genitive masculine plural}\FootnoteSubFooter\FootnoteSubHeader{d}\FnParse{present active participle genitive masculine plural}\FootnoteSubFooter\FootnoteSubHeader{e}\FnParse{present active indicative 2nd singular}\FootnoteSubFooter}\MainTextVerseMark{1}{3}ἐχάρην\MainTextFootnoteMark{a} γὰρ λίαν\MainTextFootnoteMark{b} ἐρχομένων\MainTextFootnoteMark{c} ἀδελφῶν καὶ μαρτυρούντων\MainTextFootnoteMark{d} σου τῇ ἀληθείᾳ, καθὼς σὺ ἐν ἀληθείᾳ περιπατεῖς.\MainTextFootnoteMark{e} \VerseFootnote{4}{\FootnoteSubHeader{a}\FnParseFormGloss{accusative feminine singular}{μειζότερος}{greater}\FootnoteSubFooter\FootnoteSubHeader{b}\FnParse{present active indicative 1st singular}\FootnoteSubFooter\FootnoteSubHeader{c}\FnParse{present active subjunctive 1st singular}\FootnoteSubFooter\FootnoteSubHeader{d}\FnParse{present active participle accusative neuter plural}\FootnoteSubFooter}\MainTextVerseMark{1}{4}μειζοτέραν\MainTextFootnoteMark{a} τούτων οὐκ ἔχω\MainTextFootnoteMark{b} ⸀χαράν, ἵνα ἀκούω\MainTextFootnoteMark{c} τὰ ἐμὰ τέκνα ἐν ⸀τῇ ἀληθείᾳ περιπατοῦντα.\MainTextFootnoteMark{d} \VerseFootnote{5}{\FootnoteSubHeader{a}\FnParse{present active indicative 2nd singular}\FootnoteSubFooter\FootnoteSubHeader{b}\FnParse{aorist middle subjunctive 2nd singular}\FootnoteSubFooter\FootnoteSubHeader{c}\FnParseFormGloss{accusative masculine plural}{ξένος}{strange}\FootnoteSubFooter}\MainTextVerseMark{1}{5}Ἀγαπητέ, πιστὸν ποιεῖς\MainTextFootnoteMark{a} ὃ ἐὰν ἐργάσῃ\MainTextFootnoteMark{b} εἰς τοὺς ἀδελφοὺς καὶ ⸀τοῦτο ξένους,\MainTextFootnoteMark{c} \VerseFootnote{6}{\FootnoteSubHeader{a}\FnParse{aorist active indicative 3rd plural}\FootnoteSubFooter\FootnoteSubHeader{b}\FnParse{future active indicative 2nd singular}\FootnoteSubFooter\FootnoteSubHeader{c}\FnParseFormGloss{aorist active participle nominative masculine singular}{προπέμπω}{to accompany}\FootnoteSubFooter\FootnoteSubHeader{d}\FnFormGloss{ἀξίως}{worthily}\FootnoteSubFooter}\MainTextVerseMark{1}{6}οἳ ἐμαρτύρησάν\MainTextFootnoteMark{a} σου τῇ ἀγάπῃ ἐνώπιον ἐκκλησίας, οὓς καλῶς ποιήσεις\MainTextFootnoteMark{b} προπέμψας\MainTextFootnoteMark{c} ἀξίως\MainTextFootnoteMark{d} τοῦ θεοῦ· \VerseFootnote{7}{\FootnoteSubHeader{a}\FnParse{aorist active indicative 3rd plural}\FootnoteSubFooter\FootnoteSubHeader{b}\FnParse{present active participle nominative masculine plural}\FootnoteSubFooter\FootnoteSubHeader{c}\FnParseFormGloss{genitive masculine plural}{ἐθνικός}{pagan}\FootnoteSubFooter}\MainTextVerseMark{1}{7}ὑπὲρ γὰρ τοῦ ὀνόματος ἐξῆλθον\MainTextFootnoteMark{a} μηδὲν λαμβάνοντες\MainTextFootnoteMark{b} ἀπὸ τῶν ⸀ἐθνικῶν.\MainTextFootnoteMark{c} \VerseFootnote{8}{\FootnoteSubHeader{a}\FnParse{present active indicative 1st plural}\FootnoteSubFooter\FootnoteSubHeader{b}\FnParseFormGloss{present active infinitive}{ὑπολαμβάνω}{to take up}\FootnoteSubFooter\FootnoteSubHeader{c}\FnParseFormGloss{nominative masculine plural}{συνεργός}{fellow worker}\FootnoteSubFooter\FootnoteSubHeader{d}\FnParse{present middle subjunctive 1st plural}\FootnoteSubFooter}\MainTextVerseMark{1}{8}ἡμεῖς οὖν ὀφείλομεν\MainTextFootnoteMark{a} ⸀ὑπολαμβάνειν\MainTextFootnoteMark{b} τοὺς τοιούτους, ἵνα συνεργοὶ\MainTextFootnoteMark{c} γινώμεθα\MainTextFootnoteMark{d} τῇ ἀληθείᾳ. \VerseFootnote{9}{\FootnoteSubHeader{a}\FnParse{aorist active indicative 1st singular}\FootnoteSubFooter\FootnoteSubHeader{b}\FnParseFormGloss{present active participle nominative masculine singular}{φιλοπρωτεύω}{to love to be first}\FootnoteSubFooter\FootnoteSubHeader{c}\FnParseFormGloss{nominative masculine singular}{Διοτρέφης}{Diotrephes}\FootnoteSubFooter\FootnoteSubHeader{d}\FnParseFormGloss{present middle indicative 3rd singular}{ἐπιδέχομαι}{to welcome}\FootnoteSubFooter}\MainTextVerseMark{1}{9}Ἔγραψά\MainTextFootnoteMark{a} ⸀τι τῇ ἐκκλησίᾳ· ἀλλ’ ὁ φιλοπρωτεύων\MainTextFootnoteMark{b} αὐτῶν Διοτρέφης\MainTextFootnoteMark{c} οὐκ ἐπιδέχεται\MainTextFootnoteMark{d} ἡμᾶς. \VerseFootnote{10}{\FootnoteSubHeader{a}\FnParse{aorist active subjunctive 1st singular}\FootnoteSubFooter\FootnoteSubHeader{b}\FnParseFormGloss{future active indicative 1st singular}{ὑπομιμνῄσκω}{to remind}\FootnoteSubFooter\FootnoteSubHeader{c}\FnParse{present active indicative 3rd singular}\FootnoteSubFooter\FootnoteSubHeader{d}\FnParseFormGloss{present active participle nominative masculine singular}{φλυαρέω}{to gossip}\FootnoteSubFooter\FootnoteSubHeader{e}\FnParseFormGloss{present passive participle nominative masculine singular}{ἀρκέω}{to be content}\FootnoteSubFooter\FootnoteSubHeader{f}\FnParseFormGloss{present middle indicative 3rd singular}{ἐπιδέχομαι}{to welcome}\FootnoteSubFooter\FootnoteSubHeader{g}\FnParse{present middle participle accusative masculine plural}\FootnoteSubFooter\FootnoteSubHeader{h}\FnParseFormGloss{present active indicative 3rd singular}{κωλύω}{to hinder}\FootnoteSubFooter\FootnoteSubHeader{i}\FnParse{present active indicative 3rd singular}\FootnoteSubFooter}\MainTextVerseMark{1}{10}διὰ τοῦτο, ἐὰν ἔλθω,\MainTextFootnoteMark{a} ὑπομνήσω\MainTextFootnoteMark{b} αὐτοῦ τὰ ἔργα ἃ ποιεῖ,\MainTextFootnoteMark{c} λόγοις πονηροῖς φλυαρῶν\MainTextFootnoteMark{d} ἡμᾶς, καὶ μὴ ἀρκούμενος\MainTextFootnoteMark{e} ἐπὶ τούτοις οὔτε αὐτὸς ἐπιδέχεται\MainTextFootnoteMark{f} τοὺς ἀδελφοὺς καὶ τοὺς βουλομένους\MainTextFootnoteMark{g} κωλύει\MainTextFootnoteMark{h} καὶ ἐκ τῆς ἐκκλησίας ἐκβάλλει.\MainTextFootnoteMark{i} \VerseFootnote{11}{\FootnoteSubHeader{a}\FnParseFormGloss{present middle imperative 2nd singular}{μιμέομαι}{to imitate}\FootnoteSubFooter\FootnoteSubHeader{b}\FnParseFormGloss{present active participle nominative masculine singular}{ἀγαθοποιέω}{to do good}\FootnoteSubFooter\FootnoteSubHeader{c}\FnParse{present active indicative 3rd singular}\FootnoteSubFooter\FootnoteSubHeader{d}\FnParseFormGloss{present active participle nominative masculine singular}{κακοποιέω}{to do evil}\FootnoteSubFooter\FootnoteSubHeader{e}\FnParse{perfect active indicative 3rd singular}\FootnoteSubFooter}\MainTextVerseMark{1}{11}Ἀγαπητέ, μὴ μιμοῦ\MainTextFootnoteMark{a} τὸ κακὸν ἀλλὰ τὸ ἀγαθόν. ὁ ἀγαθοποιῶν\MainTextFootnoteMark{b} ἐκ τοῦ θεοῦ ἐστιν·\MainTextFootnoteMark{c} ὁ κακοποιῶν\MainTextFootnoteMark{d} οὐχ ἑώρακεν\MainTextFootnoteMark{e} τὸν θεόν. \VerseFootnote{12}{\FootnoteSubHeader{a}\FnParseFormGloss{dative masculine singular}{Δημήτριος}{Demetrius}\FootnoteSubFooter\FootnoteSubHeader{b,c}\FnParse{perfect passive indicative 3rd singular}\FootnoteSubFooter\FootnoteSubHeader{d}\FnParse{perfect active indicative 2nd singular}\FootnoteSubFooter\FootnoteSubHeader{e}\FnParseFormGloss{nominative feminine singular}{ἀληθής}{true}\FootnoteSubFooter\FootnoteSubHeader{f}\FnParse{present active indicative 3rd singular}\FootnoteSubFooter}\MainTextVerseMark{1}{12}Δημητρίῳ\MainTextFootnoteMark{a} μεμαρτύρηται\MainTextFootnoteMark{b} ὑπὸ πάντων καὶ ὑπὸ αὐτῆς τῆς ἀληθείας· καὶ ἡμεῖς δὲ μαρτυροῦμεν,\MainTextFootnoteMark{c} καὶ ⸀οἶδας\MainTextFootnoteMark{d} ὅτι ἡ μαρτυρία ἡμῶν ἀληθής\MainTextFootnoteMark{e} ἐστιν.\MainTextFootnoteMark{f} \VerseFootnote{13}{\FootnoteSubHeader{a}\FnParse{imperfect active indicative 1st singular}\FootnoteSubFooter\FootnoteSubHeader{b,f}\FnParse{aorist active infinitive}\FootnoteSubFooter\FootnoteSubHeader{c}\FnParse{present active indicative 1st singular}\FootnoteSubFooter\FootnoteSubHeader{d}\FnParseFormGloss{genitive neuter singular}{μέλαν}{the color black}\FootnoteSubFooter\FootnoteSubHeader{e}\FnParseFormGloss{genitive masculine singular}{κάλαμος}{reed}\FootnoteSubFooter}\MainTextVerseMark{1}{13}Πολλὰ εἶχον\MainTextFootnoteMark{a} ⸂γράψαι\MainTextFootnoteMark{b} σοι⸃, ἀλλ’ οὐ θέλω\MainTextFootnoteMark{c} διὰ μέλανος\MainTextFootnoteMark{d} καὶ καλάμου\MainTextFootnoteMark{e} σοι ⸀γράφειν·\MainTextFootnoteMark{f} \VerseFootnote{14}{\FootnoteSubHeader{a}\FnParse{present active indicative 1st singular}\FootnoteSubFooter\FootnoteSubHeader{b}\FnParse{aorist active infinitive}\FootnoteSubFooter\FootnoteSubHeader{c}\FnParse{future active indicative 1st plural}\FootnoteSubFooter}\MainTextVerseMark{1}{14}ἐλπίζω\MainTextFootnoteMark{a} δὲ εὐθέως ⸂σε ἰδεῖν⸃,\MainTextFootnoteMark{b} καὶ στόμα πρὸς στόμα λαλήσομεν.\MainTextFootnoteMark{c} \VerseFootnote{15}{\FootnoteSubHeader{a,c}\FnParse{present middle indicative 3rd plural}\FootnoteSubFooter\FootnoteSubHeader{b,d}\FnParseFormGloss{nominative masculine plural}{φίλος}{friendly}\FootnoteSubFooter}\MainTextVerseMark{1}{15}Εἰρήνη σοι. ἀσπάζονταί\MainTextFootnoteMark{a} σε οἱ φίλοι.\MainTextFootnoteMark{b} ἀσπάζου\MainTextFootnoteMark{c} τοὺς φίλους\MainTextFootnoteMark{d} κατ’ ὄνομα. 
\end{document}